% ============================================================
%  REFERENCES — Complete bibliography for PairedDiff paper
%  \input'd from main.tex after \appendix
%  23 original references + [24] FLUX.1 (Black Forest Labs, 2024)
% ============================================================

\begin{thebibliography}{99}

\bibitem{wang2020cnn}
S.~Wang, O.~Wang, R.~Zhang, A.~Owens, and A.~A.~Efros,
``CNN-generated images are surprisingly easy to spot\ldots for now,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition
(CVPR)}, pp.\ 8695--8704, 2020.

\bibitem{zhu2023genimage}
M.~Zhu, H.~Chen, Q.~Yan, X.~Huang, G.~Lin, W.~Li, T.~Sun,
T.~Zhang, H.~Ji, and R.~Dong,
``GenImage: A million-scale benchmark for detecting AI-generated
image,''
in \emph{Advances in Neural Information Processing Systems (NeurIPS)},
vol.\ 36, pp.\ 77771--77782, 2023.

\bibitem{grommelt2024jpeg}
P.~Grommelt, L.~Weiss, F.~Haselbeck, and J.~N{\"o}th,
``Fake or JPEG? Revealing common biases in generated image detection
datasets,''
in \emph{Proc. Int. Conf. Computer Vision Theory and Applications
(VISAPP)}, Springer LNCS, vol.\ 15369, pp.\ 103--117, 2024.

\bibitem{hong2024wildfake}
Y.~Hong and J.~Zhang,
``WildFake: A large-scale challenging dataset for AI-generated images
detection,''
\emph{arXiv preprint arXiv:2402.11843}, 2024.

\bibitem{pellegrini2025aigenbench}
L.~Pellegrini, E.~Montagna, R.~Amoroso, G.~Boato, and
F.~De~Natale,
``AI-GenBench: A new ongoing benchmark for AI-generated image
detection,''
in \emph{Proc. Int. Joint Conf. Neural Networks (IJCNN) Workshops
(VerIMedia)}, 2025.

\bibitem{corvi2023detection}
R.~Corvi, D.~Cozzolino, G.~Zingarini, G.~Poggi, K.~Nagano, and
L.~Verdoliva,
``On the detection of synthetic images generated by diffusion
models,''
in \emph{Proc. IEEE Int. Conf. Acoustics, Speech and Signal
Processing (ICASSP)}, pp.\ 1--5, 2023.

\bibitem{jiang2024passive}
Z.~Jiang, M.~Fang, and N.~Z.~Gong,
``AI-generated image detection: Passive or watermark?''
\emph{arXiv preprint arXiv:2411.13553}, 2024.

\bibitem{li2023blip2}
J.~Li, D.~Li, S.~Savarese, and S.~Hoi,
``BLIP-2: Bootstrapping language-image pre-training with frozen image
encoders and large language models,''
in \emph{Proc. Int. Conf. Machine Learning (ICML)},
vol.\ 202, pp.\ 19730--19742, 2023.

\bibitem{rombach2022ldm}
R.~Rombach, A.~Blattmann, D.~Lorenz, P.~Esser, and B.~Ommer,
``High-resolution image synthesis with latent diffusion models,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition
(CVPR)}, pp.\ 10684--10695, 2022.

\bibitem{podell2024sdxl}
D.~Podell, Z.~English, K.~Lacey, A.~Blattmann, T.~Dockhorn,
J.~M{\"u}ller, J.~Penna, and R.~Rombach,
``SDXL: Improving latent diffusion models for high-resolution image
synthesis,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2024.

\bibitem{chen2024pixart}
J.~Chen, J.~Yu, C.~Ge, L.~Yao, E.~Xie, Y.~Wu, Z.~Wang, J.~Kwok,
P.~Luo, H.~Lu, and Z.~Liu,
``PixArt-$\Sigma$: Weak-to-strong training of diffusion transformer
for 4K text-to-image generation,''
in \emph{Proc. European Conf. Computer Vision (ECCV)},
Springer LNCS, vol.\ 15090, pp.\ 74--91, 2024.

\bibitem{razzhigaev2023kandinsky}
A.~Razzhigaev, A.~Shakhmatov, A.~Maltseva, V.~Arkhipkin,
I.~Pavlov, I.~Ryabov, A.~Kuts, A.~Panchenko, A.~Kuznetsov,
and D.~Dimitrov,
``Kandinsky: An improved text-to-image synthesis with image prior
and latent diffusion,''
\emph{arXiv preprint arXiv:2310.03502}, 2023.

\bibitem{pernias2023wuerstchen}
P.~Pernias, D.~Rampas, M.~L.~Richter, C.~Pal, and M.~Aubreville,
``W\"{u}rstchen: An efficient architecture for large-scale
text-to-image diffusion models,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2024.

\bibitem{ojha2023universal}
U.~Ojha, Y.~Li, and Y.~J.~Lee,
``Towards universal fake image detectors that generalize across
generative models,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition
(CVPR)}, pp.\ 24480--24489, 2023.

\bibitem{deng2009imagenet}
J.~Deng, W.~Dong, R.~Socher, L.-J.~Li, K.~Li, and L.~Fei-Fei,
``ImageNet: A large-scale hierarchical image database,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition
(CVPR)}, pp.\ 248--255, 2009.

\bibitem{lin2014coco}
T.-Y.~Lin, M.~Maire, S.~Belongie, J.~Hays, P.~Perona, D.~Ramanan,
P.~Doll{\'a}r, and C.~L.~Zitnick,
``Microsoft COCO: Common objects in context,''
in \emph{Proc. European Conf. Computer Vision (ECCV)},
Springer LNCS, vol.\ 8693, pp.\ 740--755, 2014.

\bibitem{radford2021clip}
A.~Radford, J.~W.~Kim, C.~Hallacy, A.~Ramesh, G.~Goh,
S.~Agarwal, G.~Sastry, A.~Askell, P.~Mishkin, J.~Clark,
G.~Krueger, and I.~Sutskever,
``Learning transferable visual models from natural language
supervision,''
in \emph{Proc. Int. Conf. Machine Learning (ICML)},
vol.\ 139, pp.\ 8748--8763, 2021.

\bibitem{he2016resnet}
K.~He, X.~Zhang, S.~Ren, and J.~Sun,
``Deep residual learning for image recognition,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition
(CVPR)}, pp.\ 770--778, 2016.

\bibitem{tan2019efficientnet}
M.~Tan and Q.~V.~Le,
``EfficientNet: Rethinking model scaling for convolutional neural
networks,''
in \emph{Proc. Int. Conf. Machine Learning (ICML)},
vol.\ 97, pp.\ 6105--6114, 2019.

\bibitem{dosovitskiy2021vit}
A.~Dosovitskiy, L.~Beyer, A.~Kolesnikov, D.~Weissenborn, X.~Zhai,
T.~Unterthiner, M.~Dehghani, M.~Minderer, G.~Heigold, S.~Gelly,
J.~Uszkoreit, and N.~Houlsby,
``An image is worth 16$\times$16 words: Transformers for image
recognition at scale,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2021.

\bibitem{loshchilov2019adamw}
I.~Loshchilov and F.~Hutter,
``Decoupled weight decay regularization,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2019.

\bibitem{gebru2021datasheets}
T.~Gebru, J.~Morgenstern, B.~Vecchione, J.~W.~Vaughan,
H.~Wallach, H.~D.~III, and K.~Crawford,
``Datasheets for datasets,''
\emph{Communications of the ACM}, vol.\ 64, no.\ 12,
pp.\ 86--92, 2021.

\bibitem{bai2026opendetect}
H.~Bai, Y.~Shi, J.~Guo, S.~Li, and X.~Wang,
``How well are open-sourced AI-generated image detection models
out-of-the-box: A comprehensive benchmark study,''
\emph{arXiv preprint arXiv:2602.07814}, 2026.

\bibitem{flux2024bfl}
Black Forest Labs,
``FLUX.1: A family of flow-matching text-to-image models,''
2024. [Online]. Available:
\url{https://github.com/black-forest-labs/flux}

\end{thebibliography}
