% Required packages: \usepackage{booktabs} \usepackage{amssymb} (for \checkmark)
% ============================================================
%  RELATED WORK + BIBLIOGRAPHY  —  NB13 Dataset Paper
%  Place \input{related_work_and_references} in your main .tex
%  Requires: \usepackage{cite}  (or biblatex with backend=biber)
%  All references in \biblio{} are cited at least once above.
% ============================================================

% ------------------------------------------------------------------
\section{Related Work}
\label{sec:related}
% ------------------------------------------------------------------

\subsection{AI-Generated Image Detection Datasets}
\label{ssec:datasets}

The problem of distinguishing synthetic from real imagery has driven the construction of several large-scale benchmarks, each making different trade-offs between scale, generator diversity, and preprocessing integrity.

Wang~\etal~\cite{wang2020cnn} introduced \textbf{ForenSynths}, the first widely adopted benchmark for this task, assembling images from eleven GAN architectures to evaluate cross-generator generalisation.
Their central finding — that a ResNet-50 classifier trained on one generator degrades sharply when transferred to another — remains one of the most influential observations in the field, and directly motivated the cross-generator evaluation protocol adopted in our work.

As generative models shifted from GANs to diffusion models, ForenSynths became inadequate.
Zhu~\etal~\cite{zhu2023genimage} addressed this with \textbf{GenImage}, a million-scale benchmark pairing real ImageNet images with synthetic counterparts from eight generators spanning both GAN and diffusion architectures.
GenImage defined two canonical evaluation tasks: cross-generator image classification and degraded image classification.
However, subsequent analysis revealed that GenImage carries a severe JPEG compression disparity: real images are sourced from ImageNet with a modal quality factor of approximately 96, whereas generated images are uncompressed PNG~\cite{grommelt2024jpeg}.
Detectors trained on GenImage thus partially function as JPEG artifact classifiers rather than true synthesis detectors, a bias that inflates reported cross-generator numbers by as much as 11 percentage points~\cite{grommelt2024jpeg}.
Our dataset eliminates this class of bias by applying a single canonical preprocessing pipeline — JPEG equalisation followed by PNG re-encoding — identically to both real and AI images, so no compression signal can leak the label.

Hong and Zhang~\cite{hong2024wildfake} subsequently introduced \textbf{WildFake}, which addresses a different gap by collecting real user-generated prompts from community platforms such as Civitai and Midjourney, thereby reflecting the diversity of prompts encountered in the wild rather than templated generation.
WildFake demonstrates that model architecture, rather than fine-tuning variant, is the primary source of generator-specific bias — a finding consistent with the per-generator detectability gradient we observe across our six generators.

The more recent \textbf{AI-GenBench}~\cite{pellegrini2025aigenbench} takes an incremental training perspective, ordering generators by historical release date and evaluating how detectors degrade as new generation methods arrive.
Their key finding is that mean detection accuracy drops from approximately 79\% for generators released in 2020--2021 to around 38\% for 2024 models, confirming an accelerating arms race.
This motivates our inclusion of recent architectures — FLUX.1-schnell, PixArt-$\Sigma$, and W\"{u}rstchen — whose detectability profiles we report as part of our baseline experiments.

\subsection{Dataset Biases and the Leak Problem}
\label{ssec:bias}

The compression bias documented in GenImage is not an isolated case.
Grommelt~\etal~\cite{grommelt2024jpeg} provide a systematic account of two pervasive biases in AI-generated image detection datasets: JPEG quality factor disparity and resolution disparity.
Because real images are typically sourced from the web in JPEG format while generated images are saved as lossless PNG, a detector has access to a trivial shortcut that is unrelated to synthesis artefacts.
Similarly, because generators produce images at fixed, model-specific resolutions (e.g., SDXL at 1024$\times$1024 versus SD~1.5 at 512$\times$512), a model can learn to key on spatial frequency statistics associated with resampling rather than generation.
Removing both biases produces measurable gains in cross-generator generalisation, which Grommelt~\etal\ attribute entirely to the correction of shortcut learning rather than to any improvement in detector architecture.

Earlier work by Corvi~\etal~\cite{corvi2023detection} characterised the forensic properties of diffusion model outputs in depth, demonstrating that existing GAN-era detectors fail almost entirely on diffusion-generated images, and that diffusion models leave qualitatively different artefacts — concentrated in specific frequency bands — compared with GAN fingerprints.
Their study used a denoising filter to isolate noise residuals, revealing model-specific spectral signatures that inform our decision to use the sniff-test AUROC as a diagnostic for per-generator artefact strength rather than as a binary pass/fail gate.

The most direct motivation for our preprocessing design comes from the caption-grounded pairing strategy proposed in ImageDetectBench~\cite{jiang2024passive}.
Jiang~\etal\ observe that if AI-generated images are produced from the same captions as their real-image counterparts, the content distribution between the two classes becomes matched, forcing detectors to rely on synthesis artefacts rather than semantic shortcuts.
Their benchmark covers four dataset--generator pairings of up to 11,000 images each.
Our dataset scales this principle to 70,000 images across six generators simultaneously, using captions produced by BLIP-2~\cite{li2023blip2} from the real images themselves, capping each prompt at 77 CLIP tokens to remain within the text encoder budget of all six generators.
Crucially, the same cleaned caption is passed to all six generators without modification, so content variation between generators is eliminated by construction.

\subsection{Generative Models Used in This Work}
\label{ssec:generators}

Our dataset draws from six text-to-image architectures that collectively span the main design families active in the period 2022--2024.

Stable Diffusion 1.5 and SDXL are latent diffusion models~\cite{rombach2022ldm,podell2024sdxl} operating in the compressed latent space of a pretrained variational autoencoder, with a UNet denoiser conditioned on text embeddings from CLIP.
SDXL extends SD~1.5 with a three-times larger UNet, a second text encoder, and native 1024$\times$1024 resolution.
Both remain the most widely deployed open-weight generators and serve as the in-distribution reference for much of the detection literature.

PixArt-$\Sigma$~\cite{chen2024pixart} replaces the UNet backbone with a Diffusion Transformer (DiT) architecture, enabling efficient generation at up to 4K resolution from a compact 0.6B parameter model.
The switch from a convolutional to an attention-based backbone produces qualitatively different artefact distributions, which is reflected in the detectability scores we report.

FLUX.1-schnell is a rectified flow transformer operating natively in latent space, representing the latest generation of open-weight text-to-image models as of mid-2024.
Its distilled, few-step inference regime produces images with fewer iterative sampling artefacts than earlier diffusion models.

Kandinsky~2.2~\cite{razzhigaev2023kandinsky} follows a two-stage prior--decoder architecture: a prior network maps text to image embeddings in CLIP space, and a latent diffusion decoder then synthesises the image.
The separation of semantic and visual decoding produces distinctive artefacts, and our sniff-test results confirm that Kandinsky images are among the most trivially detectable in a single-epoch probe.

W\"{u}rstchen~\cite{pernias2023wuerstchen} employs a hierarchical three-stage latent compression scheme inspired by the W{\"u}rstchen architecture, operating at an extreme compression ratio that distinguishes it from the single-stage latent diffusion approach.
Its one-epoch sniff-test AUROC of 1.0 indicates that its output carries the most discriminative forensic signature of all six generators in our dataset, a result we discuss in detail in Section~\ref{sec:experiments}.

\subsection{Detection Methods}
\label{ssec:detectors}

Alongside the dataset landscape, a parallel body of work has developed detector architectures.
Wang~\etal's CNNSpot~\cite{wang2020cnn} established the standard pipeline — fine-tuning a pretrained CNN (ResNet-50) as a binary real-vs-fake classifier — which, despite its simplicity, remains the dominant baseline for dataset evaluation papers including ours.

Ojha~\etal~\cite{ojha2023universal} showed that this standard pipeline fails catastrophically when trained on GAN images and evaluated on diffusion images: the real class becomes a sink that absorbs any image not matching the specific GAN artefact pattern.
Their proposed fix — using frozen CLIP features with nearest-neighbour classification rather than a learned classifier — substantially improves cross-family generalisation, motivating our inclusion of ViT-B/16 as one of our three baselines, which shares the same ViT backbone used by CLIP.

More recently, the community has explored whether the JPEG compression bias in training data is itself the reason cross-generator generalisation is poor~\cite{grommelt2024jpeg}.
The \textit{AI-generated Image Detection: Passive or Watermark?} benchmark~\cite{jiang2024passive} further argues that passive detectors are fundamentally limited by the artefacts available in their training distribution, and that watermark-based proactive detection consistently outperforms passive approaches under perturbation.
While our paper focuses on the passive detection setting — the setting most relevant to detecting images generated by models without watermarking — this comparison motivates future work using our dataset to evaluate proactive detection methods.

\subsection{Positioning of This Work}
\label{ssec:position}

Table~\ref{tab:dataset_comparison} summarises how our dataset compares against the most closely related prior benchmarks.
The key distinctions are threefold.
First, we are the only dataset to apply a \emph{single canonical preprocessing function} — identical in every detail for real and AI images — that eliminates both JPEG quality and resolution asymmetry simultaneously.
Second, our caption-grounded pairing ensures that all six generators receive byte-identical prompts derived from the same real image, eliminating semantic content as a confounding variable.
Third, we provide pair-level deterministic splits keyed on the real image identifier, guaranteeing that every split contains balanced real--AI pairs and that no image or its prompt appears in more than one fold.

\begin{table}[t]
\centering
\caption{Comparison of AI-generated image detection datasets. ``Paired'' denotes that real and AI images share a common content anchor (caption or class). ``Leak-free'' denotes that an identical preprocessing pipeline was applied to both classes.}
\label{tab:dataset_comparison}
\begin{tabular}{lccccl}
\toprule
Dataset & Images & Generators & Paired & Leak-free & Venue \\
\midrule
ForenSynths~\cite{wang2020cnn}        & 362K  & 11 (GANs)        & $\times$ & $\times$ & CVPR 2020 \\
GenImage~\cite{zhu2023genimage}       & 1.3M  & 8                & $\times$ & $\times$ & NeurIPS 2023 \\
WildFake~\cite{hong2024wildfake}      & 84K   & 17+              & $\times$ & $\times$ & arXiv 2024 \\
AI-GenBench~\cite{pellegrini2025aigenbench} & 310K & 20+         & $\times$ & $\times$ & IJCNN 2025 \\
ImageDetectBench~\cite{jiang2024passive} & 44K & 4              & \checkmark & $\times$ & arXiv 2024 \\
\midrule
\textbf{Ours}                         & \textbf{70K} & \textbf{6} & \checkmark & \checkmark & -- \\
\bottomrule
\end{tabular}
\end{table}

% ------------------------------------------------------------------
% BIBLIOGRAPHY
% All 14 entries are cited at least once in Related Work above.
% ------------------------------------------------------------------

\begin{thebibliography}{99}

\bibitem{wang2020cnn}
S.~Wang, O.~Wang, R.~Zhang, A.~Owens, and A.~A.~Efros,
``CNN-generated images are surprisingly easy to spot\ldots for now,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition (CVPR)},
pp.\ 8695--8704, 2020.

\bibitem{zhu2023genimage}
M.~Zhu, H.~Chen, Q.~Yan, X.~Huang, G.~Lin, W.~Li, Z.~Tu, H.~Hu, J.~Hu, and Y.~Wang,
``GenImage: A million-scale benchmark for detecting AI-generated image,''
in \emph{Advances in Neural Information Processing Systems (NeurIPS)},
vol.\ 36, pp.\ 77771--77782, 2023.

\bibitem{grommelt2024jpeg}
P.~Grommelt, L.~Weiss, F.~Haselbeck, and J.~Nöth,
``Fake or JPEG? Revealing common biases in generated image detection datasets,''
in \emph{Proc. Int. Conf. Computer Vision Theory and Applications (VISAPP)},
Springer LNCS, vol.\ 15369, pp.\ 103--117, 2024.

\bibitem{hong2024wildfake}
Y.~Hong and J.~Zhang,
``WildFake: A large-scale challenging dataset for AI-generated images detection,''
\emph{arXiv preprint arXiv:2402.11843}, 2024.

\bibitem{pellegrini2025aigenbench}
L.~Pellegrini, D.~Cozzolino, S.~Pandolfini, D.~Maltoni, M.~Ferrara,
L.~Verdoliva, M.~Prati, and M.~Ramilli,
``AI-GenBench: A new ongoing benchmark for AI-generated image detection,''
in \emph{Proc. Int. Joint Conf. Neural Networks (IJCNN) Workshops (VerIMedia)},
2025.

\bibitem{corvi2023detection}
R.~Corvi, D.~Cozzolino, G.~Zingarini, G.~Poggi, K.~Nagano, and L.~Verdoliva,
``On the detection of synthetic images generated by diffusion models,''
in \emph{Proc. IEEE Int. Conf. Acoustics, Speech and Signal Processing (ICASSP)},
pp.\ 1--5, 2023.

\bibitem{jiang2024passive}
Z.~Jiang, M.~Fang, and N.~Z.~Gong,
``AI-generated image detection: Passive or watermark?''
\emph{arXiv preprint arXiv:2411.13553}, 2024.

\bibitem{li2023blip2}
J.~Li, D.~Li, S.~Savarese, and S.~Hoi,
``BLIP-2: Bootstrapping language-image pre-training with frozen image encoders
and large language models,''
in \emph{Proc. Int. Conf. Machine Learning (ICML)},
vol.\ 202, pp.\ 19730--19742, 2023.

\bibitem{rombach2022ldm}
R.~Rombach, A.~Blattmann, D.~Lorenz, P.~Esser, and B.~Ommer,
``High-resolution image synthesis with latent diffusion models,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition (CVPR)},
pp.\ 10684--10695, 2022.

\bibitem{podell2024sdxl}
D.~Podell, Z.~English, K.~Lacey, A.~Blattmann, T.~Dockhorn, J.~Müller,
J.~Penna, and R.~Rombach,
``SDXL: Improving latent diffusion models for high-resolution image synthesis,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2024.

\bibitem{chen2024pixart}
J.~Chen, C.~Ge, E.~Xie, Y.~Wu, L.~Yao, X.~Ren, Z.~Wang, P.~Luo, H.~Lu,
and Z.~Li,
``PixArt-$\Sigma$: Weak-to-strong training of diffusion transformer
for 4K text-to-image generation,''
in \emph{Proc. European Conf. Computer Vision (ECCV)},
Springer LNCS, vol.\ 15090, pp.\ 74--91, 2024.

\bibitem{razzhigaev2023kandinsky}
A.~Razzhigaev, A.~Shakhmatov, A.~Maltseva, V.~Arkhipkin, I.~Pavlov,
I.~Ryabov, A.~Kuts, A.~Panchenko, A.~Dimitrov, and D.~Dimitrov,
``Kandinsky: An improved text-to-image synthesis with image prior and latent diffusion,''
\emph{arXiv preprint arXiv:2310.03502}, 2023.

\bibitem{pernias2023wuerstchen}
P.~Pernias, D.~Rampas, M.~L.~Richter, C.~Pal, and M.~Aubreville,
``W\"{u}rstchen: An efficient architecture for large-scale
text-to-image diffusion models,''
in \emph{Proc. Int. Conf. Learning Representations (ICLR)}, 2024.

\bibitem{ojha2023universal}
U.~Ojha, Y.~Li, and Y.~J.~Lee,
``Towards universal fake image detectors that generalize across generative models,''
in \emph{Proc. IEEE/CVF Conf. Computer Vision and Pattern Recognition (CVPR)},
pp.\ 24480--24489, 2023.

\end{thebibliography}
